---
title: "RepData_Project1"
author: "Michelle Muha"
date: "May 5, 2015"
output: html_document
---

##Load dataset and libraries- 
```{r}
activity <- read.csv("/Users/michelle/activity.csv")
library(dplyr)
library(ggplot2)
```

##Total number of steps per day -
```{r}
data <- summarise(group_by(activity, date), sum(steps))
names(data) <- c("date", "sum")
hist(data$sum, main = "", xlab = "Total steps per day")
```

##Mean steps per day -
```{r}
mean(data$sum, na.rm = TRUE)
```

##Median steps per day -
```{r}
median(data$sum, na.rm = TRUE)
```

##Daily activity pattern - 
```{r}
data2 <- summarise(group_by(activity, interval), mean(steps, na.rm = TRUE))
names(data2) <- c("interval", "steps")
plot(data2$interval, data2$steps, xlab="Interval", ylab = "Steps", type = "l")
```

##Interval with maximum average number of steps - 
```{r}
data2$interval[which(data2$steps == max(data2$steps))]
```

##Number of missing data points -
```{r}
logic <- is.na(activity$steps)
missing <- logic[logic==TRUE]
length(missing)
```

##Adjusted dataset -
Missing data points were filled in with the average for the given interval over the dataset. Since data points were missing in chunks from entire days, the impact of the added data points was minimal
```{r}
activity_fill<-activity
for (i in 1:17568) {
  if (is.na(activity_fill$steps[i]) ==  TRUE) {
    interval <- activity_fill$interval[i]
    index <- which(data2$interval == interval)
    mean <- data2$steps[index]
    activity_fill$steps[i] <- mean 
  }
}
```

##Number of steps per day (adjusted) -
```{r}
data <- summarise(group_by(activity_fill, date), sum(steps))
names(data) <- c("date", "sum")
hist(data$sum, main = "", xlab = "Total steps per day")
```

##Mean steps per day (adjusted)-
```{r}
mean(data$sum, na.rm = TRUE)
```

##Median steps per day (adjusted)-
```{r}
median(data$sum, na.rm = TRUE)
```

##Differences between weekdays / weekends
```{r}
activity_fill <- mutate(activity_fill, day = weekdays(as.Date(as.character(date))))
activity_fill <- mutate(activity_fill, weekend = day %in% c("Saturday", "Sunday"))
data3 <- summarise(group_by(activity_fill, weekend, interval), mean(steps))
names(data3) <- c("weekday", "interval", "mean.steps")
weekdays <- data3[data3$weekday == "TRUE",]
weekends <- data3[data3$weekday == "FALSE",]
par(mfcol=c(2,1))
plot(weekdays$interval, weekdays$mean.steps, xlab="Interval", ylab = "Steps", type = "l", main = "Weekdays")
plot(weekends$interval, weekends$mean.steps, xlab="Interval", ylab = "Steps", type = "l", main = "Weekends")
```